---
title: "ESG Tourism — Reviewer Diagnostics"
subtitle: "Within R², SEs, sample, COVID robustness, v2-vs-v3"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
    code-summary: "Show code"
    self-contained: true
    fig-width: 10
    fig-height: 6
execute:
  echo: true
  warning: false
  message: false
---

# Setup

```{r setup}
#| code-fold: true
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(plm)
  library(lmtest)
  library(sandwich)
  library(knitr)
  library(kableExtra)
})

options(scipen = 999)

dta_clean <- readRDS("dta_clean.rds")

# Recompute developed/dev_group exactly as in v3 (median of country-mean GDP)
country_gdp <- dta_clean %>%
  group_by(Country) %>%
  summarise(avg_gdp = mean(gov_gdp_capita, na.rm = TRUE), .groups = "drop")

median_gdp <- median(country_gdp$avg_gdp, na.rm = TRUE)
country_gdp <- country_gdp %>%
  mutate(developed = as.integer(avg_gdp > median_gdp),
         dev_group = ifelse(developed == 1, "High Income", "Low/Middle Income"))

dta_clean <- dta_clean %>%
  left_join(country_gdp %>% select(Country, developed, dev_group), by = "Country")

pdata <- pdata.frame(dta_clean, index = c("Country", "Year"), drop.index = FALSE)

# Spec definitions
v3_ivs <- c(
  "gov_corruption", "gov_gdp_capita", "gov_effectiveness", "gov_pol_stability", "gov_rule_law",
  "soc_internet_users", "soc_life_expect", "soc_unempl_ilo", "soc_broadband", "soc_health_exp", "soc_mort_infant",
  "env_pm25_exp", "env_co2_capita", "env_renew_elec", "env_nd_gain_idx"
)
v3_ivs <- v3_ivs[v3_ivs %in% names(pdata)]

v2_ivs <- c(
  "gov_corruption", "soc_internet_users", "env_co2_capita",
  "env_renew_elec", "env_pm25_exp", "soc_unempl_ilo",
  "gov_gdp_capita", "soc_life_expect"
)
v2_ivs <- v2_ivs[v2_ivs %in% names(pdata)]

dvs <- c("log_arrivals", "log_nights", "log_night_pop", "log_night_area")
dvs <- dvs[dvs %in% names(pdata)]

key4 <- c("gov_corruption", "env_pm25_exp", "soc_internet_users", "soc_unempl_ilo")
```

```{r helpers}
#| code-fold: true

# Two-way FE with HC1 cluster-by-country SE; returns coeftest, fitted plm, within R²
fit_twfe <- function(dv, ivs, data) {
  f <- as.formula(paste(dv, "~", paste(ivs, collapse = " + ")))
  fe2 <- plm(f, data = data, model = "within", effect = "twoways",
             index = c("Country", "Year"))
  ct  <- coeftest(fe2, vcov = vcovHC(fe2, type = "HC1", cluster = "group"))
  list(model = fe2,
       ct = ct,
       r2 = unname(summary(fe2)$r.squared["rsq"]),
       adj_r2 = unname(summary(fe2)$r.squared["adjrsq"]),
       n = nobs(fe2),
       n_countries = length(unique(index(fe2)[[1]])))
}

# Coefficient table with SE, t, p, sig stars, 95% CI
coef_kable <- function(fit, caption, dv = NULL) {
  m  <- as.matrix(fit$ct)
  df <- data.frame(
    Variable = rownames(m),
    Coefficient = round(m[, 1], 4),
    SE = round(m[, 2], 4),
    t = round(m[, 3], 2),
    p = round(m[, 4], 3),
    CI_low  = round(m[, 1] - 1.96 * m[, 2], 4),
    CI_high = round(m[, 1] + 1.96 * m[, 2], 4),
    row.names = NULL
  )
  df$Sig <- with(df, ifelse(p < 0.01, "***", ifelse(p < 0.05, "**", ifelse(p < 0.10, "*", ""))))
  df <- df[, c("Variable", "Coefficient", "SE", "t", "p", "Sig", "CI_low", "CI_high")]

  footer <- sprintf("n = %d | countries = %d | within R² = %.3f | adj. within R² = %.3f",
                    fit$n, fit$n_countries, fit$r2, fit$adj_r2)
  cap <- if (is.null(dv)) caption else paste0(caption, " — DV: ", dv)
  kable(df, caption = paste0(cap, "  •  ", footer)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

# Headline coefficient tables — full sample, v3 spec

These are the v3 two-way FE results with cluster-robust HC1 SEs, **with within R² printed under every table**. Apply the user's rule of thumb to the within R²: <0.05 → noise; 0.05–0.15 → ambiguous; >0.15 → real.

```{r headline-full}
#| code-fold: true
#| results: asis

for (dv in dvs) {
  fit <- fit_twfe(dv, v3_ivs, pdata)
  print(coef_kable(fit, "Two-way FE (full sample, v3 spec)", dv))
  cat("\n\n")
}
```

# Country list with GDP-per-capita classification

The v3 paper splits countries at the **median of country-mean `gov_gdp_capita` across 2000–2023** — *not* a World Bank or IMF classification. The threshold and the actual country bins are below; the user can confirm whether the "developing" label is defensible.

```{r country-classification}
#| code-fold: true

country_summary <- dta_clean %>%
  group_by(Country) %>%
  summarise(
    avg_gdp_pc = round(mean(gov_gdp_capita, na.rm = TRUE), 0),
    min_year = min(Year, na.rm = TRUE),
    max_year = max(Year, na.rm = TRUE),
    n_years = n(),
    .groups = "drop"
  ) %>%
  mutate(
    threshold = round(median_gdp, 0),
    distance_from_median = round(avg_gdp_pc - median_gdp, 0),
    group = ifelse(avg_gdp_pc > median_gdp, "High Income", "Low/Middle Income")
  ) %>%
  arrange(desc(avg_gdp_pc)) %>%
  select(Country, avg_gdp_pc, threshold, distance_from_median, group, min_year, max_year, n_years)

kable(country_summary,
      col.names = c("Country", "Avg GDP/cap", "Median threshold",
                    "Δ from median", "Group", "Start", "End", "N years"),
      caption = sprintf("Country classification (median = %s). 13 countries above, 14 below — or vice-versa, depending on ties.",
                        format(round(median_gdp, 0), big.mark = ","))) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  row_spec(which(country_summary$group == "Low/Middle Income"), background = "#fff5f5")
```

# Within-country variation in the four key regressors

If a country has only a handful of distinct values for `gov_corruption` across 24 years, the within-country FE estimator has very little variation to identify the coefficient. This section shows how much actual variation drives the FE estimates.

## Per-variable summary across the 27 countries

```{r within-summary}
#| code-fold: true

within_stats <- dta_clean %>%
  select(Country, all_of(key4)) %>%
  pivot_longer(-Country, names_to = "variable", values_to = "value") %>%
  group_by(variable, Country) %>%
  summarise(
    n_distinct = n_distinct(round(value, 6)),
    sd_within = sd(value, na.rm = TRUE),
    range_within = max(value, na.rm = TRUE) - min(value, na.rm = TRUE),
    n_obs = sum(!is.na(value)),
    .groups = "drop"
  )

within_summary <- within_stats %>%
  group_by(variable) %>%
  summarise(
    `min n_distinct` = min(n_distinct),
    `median n_distinct` = median(n_distinct),
    `max n_distinct` = max(n_distinct),
    `min within SD` = round(min(sd_within, na.rm = TRUE), 3),
    `median within SD` = round(median(sd_within, na.rm = TRUE), 3),
    `max within SD` = round(max(sd_within, na.rm = TRUE), 3),
    .groups = "drop"
  )

kable(within_summary,
      caption = "Within-country variation in key regressors (across 27 countries, 2000–2023)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Per-country detail — distinct values

```{r within-per-country}
#| code-fold: true

per_country_distinct <- within_stats %>%
  select(Country, variable, n_distinct) %>%
  pivot_wider(names_from = variable, values_from = n_distinct) %>%
  arrange(Country)

kable(per_country_distinct,
      caption = "Number of distinct within-country values, 2000–2023") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Per-country detail — within-country SD

```{r within-per-country-sd}
#| code-fold: true

per_country_sd <- within_stats %>%
  mutate(sd_within = round(sd_within, 3)) %>%
  select(Country, variable, sd_within) %>%
  pivot_wider(names_from = variable, values_from = sd_within) %>%
  arrange(Country)

kable(per_country_sd,
      caption = "Within-country standard deviation, 2000–2023") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Headline FE — with vs without 2020–2022 (COVID)

If excluding 2020–2022 materially moves a coefficient, the COVID period was doing the heavy lifting. Threshold to flag: |Δ| > 50% of full-sample β, or a sign flip, or a significance class change.

```{r covid-comparison}
#| code-fold: true
#| results: asis

pdata_no_covid <- pdata.frame(
  dta_clean %>% filter(!Year %in% 2020:2022),
  index = c("Country", "Year"), drop.index = FALSE
)

build_compare <- function(dv) {
  full <- fit_twfe(dv, v3_ivs, pdata)
  noco <- fit_twfe(dv, v3_ivs, pdata_no_covid)

  m_full <- as.matrix(full$ct)
  m_noco <- as.matrix(noco$ct)

  vars <- union(rownames(m_full), rownames(m_noco))
  cmp <- data.frame(
    Variable = vars,
    `Full β`        = round(m_full[vars, 1], 4),
    `Full SE`       = round(m_full[vars, 2], 4),
    `Full p`        = round(m_full[vars, 4], 3),
    `No-COVID β`    = round(m_noco[vars, 1], 4),
    `No-COVID SE`   = round(m_noco[vars, 2], 4),
    `No-COVID p`    = round(m_noco[vars, 4], 3),
    check.names = FALSE,
    row.names = NULL
  )
  cmp$`Δ (No-COVID − Full)` <- round(cmp$`No-COVID β` - cmp$`Full β`, 4)
  cmp$`|Δ| / |Full β|` <- ifelse(abs(cmp$`Full β`) > 0,
                                  round(abs(cmp$`Δ (No-COVID − Full)`) / abs(cmp$`Full β`), 2),
                                  NA)

  footer <- sprintf("Full: n=%d, within R²=%.3f  |  No-COVID: n=%d, within R²=%.3f",
                    full$n, full$r2, noco$n, noco$r2)
  list(table = cmp, footer = footer)
}

for (dv in dvs) {
  out <- build_compare(dv)
  print(
    kable(out$table,
          caption = paste0("Full sample vs 2020–2022 dropped — DV: ", dv, "  •  ", out$footer)) %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
  )
  cat("\n\n")
}
```

# v2 vs v3 main estimates — same data, same FE, different regressor set

The two versions of the paper use the same panel, same DV, same two-way FE, and same cluster-robust SEs — but the regressor set differs. Below, **both specs are estimated on the same `pdata`**, so any difference in `gov_corruption` etc. is *purely* attributable to omitted-variable / multicollinearity reshuffling between specifications. This is the "moving target" check.

- **v2 spec (8 IVs)**: `r paste(v2_ivs, collapse = ", ")`
- **v3 spec (15 IVs)**: `r paste(v3_ivs, collapse = ", ")`

```{r v2-vs-v3}
#| code-fold: true

fit_v2 <- fit_twfe("log_arrivals", v2_ivs, pdata)
fit_v3 <- fit_twfe("log_arrivals", v3_ivs, pdata)

m_v2 <- as.matrix(fit_v2$ct)
m_v3 <- as.matrix(fit_v3$ct)

all_vars <- union(rownames(m_v2), rownames(m_v3))

fmt <- function(M, v) {
  if (!v %in% rownames(M)) return(c(NA_real_, NA_real_, NA_real_))
  c(round(M[v, 1], 4), round(M[v, 2], 4), round(M[v, 4], 3))
}

cmp <- do.call(rbind, lapply(all_vars, function(v) {
  a <- fmt(m_v2, v); b <- fmt(m_v3, v)
  data.frame(
    Variable = v,
    `v2 β` = a[1], `v2 SE` = a[2], `v2 p` = a[3],
    `v3 β` = b[1], `v3 SE` = b[2], `v3 p` = b[3],
    `In v2` = v %in% rownames(m_v2),
    `In v3` = v %in% rownames(m_v3),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}))

cmp$Status <- with(cmp,
  ifelse(`In v2` & `In v3`, "Both",
  ifelse(`In v2`, "v2 only", "v3 only")))

# sign flip / significance class flag
sig_class <- function(p) ifelse(is.na(p), "na",
                                ifelse(p < 0.01, "***",
                                ifelse(p < 0.05, "**",
                                ifelse(p < 0.10, "*", "ns"))))
cmp$Flag <- with(cmp, ifelse(
  Status == "Both" & !is.na(`v2 β`) & !is.na(`v3 β`) &
    sign(`v2 β`) != sign(`v3 β`), "SIGN FLIP",
  ifelse(Status == "Both" & sig_class(`v2 p`) != sig_class(`v3 p`), "sig-class change", "")))

cmp <- cmp %>%
  mutate(`In v2` = ifelse(`In v2`, "yes", "no"),
         `In v3` = ifelse(`In v3`, "yes", "no")) %>%
  arrange(desc(Status == "Both"), Variable)

footer <- sprintf("v2: n=%d, within R²=%.3f, p=%d regressors  |  v3: n=%d, within R²=%.3f, p=%d regressors",
                  fit_v2$n, fit_v2$r2, length(v2_ivs),
                  fit_v3$n, fit_v3$r2, length(v3_ivs))

kable(cmp, caption = paste0("v2 vs v3 — DV: log_arrivals, two-way FE, HC1 cluster-by-country  •  ", footer)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Heterogeneity — within R² for developed and developing samples

```{r heterogeneity-r2}
#| code-fold: true

pdata_dev_sub  <- pdata[pdata$developed == 1, ]
pdata_devg_sub <- pdata[pdata$developed == 0, ]

fit_dev_full  <- fit_twfe("log_arrivals", v3_ivs, pdata_dev_sub)
fit_devg_full <- fit_twfe("log_arrivals", v3_ivs, pdata_devg_sub)

# Bivariate refit on gov_corruption alone, by subsample
fit_dev_corr  <- fit_twfe("log_arrivals", "gov_corruption", pdata_dev_sub)
fit_devg_corr <- fit_twfe("log_arrivals", "gov_corruption", pdata_devg_sub)

corr_row <- function(fit) {
  m <- as.matrix(fit$ct)
  if (!"gov_corruption" %in% rownames(m)) return("—")
  sprintf("%.4f (%.4f)%s",
          m["gov_corruption", 1], m["gov_corruption", 2],
          ifelse(m["gov_corruption", 4] < 0.01, "***",
          ifelse(m["gov_corruption", 4] < 0.05, "**",
          ifelse(m["gov_corruption", 4] < 0.10, "*", ""))))
}

het_tab <- data.frame(
  Sample = c("High Income (developed)", "Low/Middle Income (developing)"),
  n = c(fit_dev_full$n, fit_devg_full$n),
  `n countries` = c(fit_dev_full$n_countries, fit_devg_full$n_countries),
  `Within R² (full v3 spec)` = round(c(fit_dev_full$r2, fit_devg_full$r2), 3),
  `Within R² (corruption only)` = round(c(fit_dev_corr$r2, fit_devg_corr$r2), 3),
  `gov_corruption β (SE) — full spec` = c(corr_row(fit_dev_full),  corr_row(fit_devg_full)),
  `gov_corruption β (SE) — bivariate` = c(corr_row(fit_dev_corr),  corr_row(fit_devg_corr)),
  check.names = FALSE
)

kable(het_tab,
      caption = "Heterogeneity — full-spec within R² and corruption-only within R² by subsample (DV: log_arrivals, two-way FE, HC1 cluster-by-country)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Governance-pillar collinearity — what killed `gov_corruption`

The five governance variables in v3 are conceptually overlapping (Worldwide Governance Indicators are by construction correlated). This section quantifies it: pairwise correlations on the within-country-demeaned governance variables (the variation actually used by the FE estimator), and VIFs from a within-transformed regression.

```{r gov-collinearity}
#| code-fold: true

gov_pillar <- c("gov_corruption", "gov_effectiveness", "gov_pol_stability",
                "gov_rule_law", "gov_gdp_capita")

# Within-country demeaned values (the variation FE uses)
demean_within <- function(df, vars) {
  df %>%
    group_by(Country) %>%
    mutate(across(all_of(vars), ~ .x - mean(.x, na.rm = TRUE), .names = "{.col}__w")) %>%
    ungroup() %>%
    select(ends_with("__w"))
}

gov_w <- demean_within(dta_clean, gov_pillar)
names(gov_w) <- sub("__w$", "", names(gov_w))

cor_mat <- round(cor(gov_w, use = "pairwise.complete.obs"), 3)

kable(cor_mat,
      caption = "Within-country (FE-relevant) correlation matrix — governance pillar") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r gov-vif}
#| code-fold: true

# VIF from within-transformed OLS: regress each gov var on the other gov vars
# using within-demeaned values, so VIFs reflect what the FE estimator sees.
vif_table <- data.frame(
  Variable = gov_pillar,
  VIF = NA_real_,
  `R² of aux. regression` = NA_real_,
  check.names = FALSE
)
for (i in seq_along(gov_pillar)) {
  v <- gov_pillar[i]
  others <- setdiff(gov_pillar, v)
  d <- gov_w
  d <- d[complete.cases(d), ]
  aux <- lm(as.formula(paste0(v, " ~ ", paste(others, collapse = " + "))), data = d)
  r2 <- summary(aux)$r.squared
  vif_table$`R² of aux. regression`[i] <- round(r2, 3)
  vif_table$VIF[i] <- round(1 / (1 - r2), 2)
}

kable(vif_table,
      caption = "VIF on within-country-demeaned governance variables (rule of thumb: VIF > 5 problematic, > 10 severe)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r gov-leave-one-in}
#| code-fold: true

# Sequential refit: corruption alone, then add gov vars one at a time.
# Shows how each governance variable sucks out corruption's coefficient.
add_order <- c("gov_corruption", "gov_gdp_capita", "gov_rule_law",
               "gov_effectiveness", "gov_pol_stability")

seq_rows <- list()
for (k in seq_along(add_order)) {
  ivs_k <- add_order[1:k]
  fit_k <- fit_twfe("log_arrivals", ivs_k, pdata)
  m <- as.matrix(fit_k$ct)
  if ("gov_corruption" %in% rownames(m)) {
    seq_rows[[k]] <- data.frame(
      `Step` = k,
      `Spec` = paste(ivs_k, collapse = " + "),
      `corruption β` = round(m["gov_corruption", 1], 4),
      `corruption SE` = round(m["gov_corruption", 2], 4),
      `corruption p` = round(m["gov_corruption", 4], 3),
      `Within R²` = round(fit_k$r2, 3),
      check.names = FALSE,
      stringsAsFactors = FALSE
    )
  }
}
seq_tab <- do.call(rbind, seq_rows)

kable(seq_tab,
      caption = "Sequential governance build-up: how each added control changes the corruption coefficient") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Heterogeneity — PM2.5 and internet, by income bin, COVID excluded

Re-estimates the **full v3 multivariate spec** on the developed and developing subsamples with 2020–2022 dropped, and reports the two coefficients the headline narrative depends on: `env_pm25_exp` and `soc_internet_users`. If both are noise in at least one bin, the heterogeneity story for PM2.5 / internet does not survive.

```{r het-pm25-internet}
#| code-fold: true

pdata_dev_noco  <- pdata.frame(
  dta_clean %>% filter(developed == 1, !Year %in% 2020:2022),
  index = c("Country", "Year"), drop.index = FALSE
)
pdata_devg_noco <- pdata.frame(
  dta_clean %>% filter(developed == 0, !Year %in% 2020:2022),
  index = c("Country", "Year"), drop.index = FALSE
)

fit_dev_noco  <- fit_twfe("log_arrivals", v3_ivs, pdata_dev_noco)
fit_devg_noco <- fit_twfe("log_arrivals", v3_ivs, pdata_devg_noco)

extract_var_row <- function(fit, varname) {
  m <- as.matrix(fit$ct)
  if (!varname %in% rownames(m)) return(rep(NA_real_, 4))
  c(m[varname, 1], m[varname, 2], m[varname, 3], m[varname, 4])
}

mk_row <- function(label, fit, varname) {
  v <- extract_var_row(fit, varname)
  sig <- if (is.na(v[4])) "" else if (v[4] < 0.01) "***" else if (v[4] < 0.05) "**" else if (v[4] < 0.10) "*" else ""
  data.frame(
    Sample = label,
    Variable = varname,
    β = round(v[1], 4),
    SE = round(v[2], 4),
    t = round(v[3], 2),
    p = round(v[4], 3),
    Sig = sig,
    `CI low` = round(v[1] - 1.96 * v[2], 4),
    `CI high` = round(v[1] + 1.96 * v[2], 4),
    n = fit$n,
    `Within R²` = round(fit$r2, 3),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}

het_no_covid <- rbind(
  mk_row("High Income, no COVID",   fit_dev_noco,  "env_pm25_exp"),
  mk_row("Low/Middle, no COVID",    fit_devg_noco, "env_pm25_exp"),
  mk_row("High Income, no COVID",   fit_dev_noco,  "soc_internet_users"),
  mk_row("Low/Middle, no COVID",    fit_devg_noco, "soc_internet_users")
)

kable(het_no_covid,
      caption = "PM2.5 and internet — full v3 spec, by income bin, 2020–2022 excluded (DV: log_arrivals, two-way FE, HC1 cluster-by-country)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Theory-driven sparse spec

A 5-regressor specification chosen on substantive grounds rather than statistical selection. Selection criteria:

- **`gov_corruption`** — the headline governance signal the paper wants to make. Replaces the WGI cluster (which is multi-collinear) with the single best-known indicator.
- **`env_pm25_exp`** — direct, interpretable environmental amenity (visible air quality). Substantive prior for tourism.
- **`soc_internet_users`** — booking, search, planning channel. Direct mechanism for tourism flows.
- **`soc_unempl_ilo`** — domestic-economy proxy; controls for cyclical labour-market conditions tourists are sensitive to.
- **`gov_gdp_capita`** — income / cost-of-destination control; without it, PM2.5 and internet pick up trends in development.

Five regressors at 27 countries × 24 years gives a comfortable d.f. budget and avoids the WGI collinearity problem.

```{r theory-sparse}
#| code-fold: true

theory_ivs <- c("gov_corruption", "env_pm25_exp", "soc_internet_users",
                "soc_unempl_ilo", "gov_gdp_capita")

fit_theory_full   <- fit_twfe("log_arrivals", theory_ivs, pdata)
fit_theory_noco   <- fit_twfe("log_arrivals", theory_ivs, pdata.frame(
  dta_clean %>% filter(!Year %in% 2020:2022),
  index = c("Country", "Year"), drop.index = FALSE))
fit_theory_dev    <- fit_twfe("log_arrivals", theory_ivs, pdata_dev_sub)
fit_theory_devg   <- fit_twfe("log_arrivals", theory_ivs, pdata_devg_sub)

print(coef_kable(fit_theory_full, "Theory-driven sparse spec, full sample", "log_arrivals"))
print(coef_kable(fit_theory_noco, "Theory-driven sparse spec, COVID excluded", "log_arrivals"))
```

## Theory-sparse vs v3 — side-by-side on the four headline regressors

```{r theory-vs-v3}
#| code-fold: true

four <- c("gov_corruption", "env_pm25_exp", "soc_internet_users", "soc_unempl_ilo")

side_row <- function(varname) {
  get_one <- function(fit) {
    m <- as.matrix(fit$ct)
    if (!varname %in% rownames(m)) return(rep(NA_real_, 3))
    c(m[varname, 1], m[varname, 2], m[varname, 4])
  }
  v3_full   <- get_one(fit_v3)
  th_full   <- get_one(fit_theory_full)
  th_noco   <- get_one(fit_theory_noco)
  th_dev    <- get_one(fit_theory_dev)
  th_devg   <- get_one(fit_theory_devg)

  data.frame(
    Variable = varname,
    `v3 β (SE, p)`        = sprintf("%.4f (%.4f, %.3f)", v3_full[1],  v3_full[2],  v3_full[3]),
    `Theory β (SE, p)`    = sprintf("%.4f (%.4f, %.3f)", th_full[1],  th_full[2],  th_full[3]),
    `Theory no-COVID`     = sprintf("%.4f (%.4f, %.3f)", th_noco[1],  th_noco[2],  th_noco[3]),
    `Theory High-Income`  = sprintf("%.4f (%.4f, %.3f)", th_dev[1],   th_dev[2],   th_dev[3]),
    `Theory Low/Middle`   = sprintf("%.4f (%.4f, %.3f)", th_devg[1],  th_devg[2],  th_devg[3]),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}

side_tab <- do.call(rbind, lapply(four, side_row))

footer <- sprintf("v3 full: n=%d, R²=%.3f  |  Theory full: n=%d, R²=%.3f  |  Theory no-COVID: n=%d, R²=%.3f  |  Theory High: n=%d, R²=%.3f  |  Theory Low: n=%d, R²=%.3f",
                  fit_v3$n, fit_v3$r2,
                  fit_theory_full$n, fit_theory_full$r2,
                  fit_theory_noco$n, fit_theory_noco$r2,
                  fit_theory_dev$n,  fit_theory_dev$r2,
                  fit_theory_devg$n, fit_theory_devg$r2)

kable(side_tab,
      caption = paste0("Theory-driven 5-regressor spec vs v3 — DV: log_arrivals, two-way FE, HC1 cluster-by-country  •  ", footer)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Quick-read summary

§7 shows what killed `gov_corruption` — the WGI governance variables are mechanically collinear, so the corruption coefficient is being absorbed by `gov_rule_law` / `gov_effectiveness` rather than genuinely zero. §8 decides whether the PM2.5 / internet heterogeneity story survives once both COVID and the multivariate-FE-noise are controlled. §9 gives a defensible final spec: if the theory-sparse coefficients on PM2.5 and internet are close to v3's, that is the spec to put in the main paper, with v3 demoted to an appendix robustness panel.
