---
title: "ESG Factors and International Tourism"
subtitle: "A Multi-Method Panel Analysis of 27 Countries (2000-2023)"
author: "[Your Name]"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
    code-summary: "Show code"
    self-contained: true
    fig-width: 10
    fig-height: 6
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
#| code-fold: true
#| code-summary: "Load packages and setup"

# =============================================================================
# LOAD REQUIRED PACKAGES
# =============================================================================

# Load MASS first to avoid select() conflict
if(!require("MASS")) install.packages("MASS", repos = "https://cloud.r-project.org")
library(MASS)

required_packages <- c(
  "tidyverse",
  "readxl",
  "plm",
  "lmtest",
  "sandwich",
  "randomForest",
  "glmnet",
  "zoo",
  "broom",
  "knitr",
  "kableExtra"
)

# Install missing packages
new_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages, repos = "https://cloud.r-project.org")

# Load packages
invisible(lapply(required_packages, library, character.only = TRUE))

# Ensure dplyr's select is used
select <- dplyr::select

# Set options
options(scipen = 999)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Executive Summary {.unnumbered}

::: {.callout-note appearance="simple"}
## Key Findings

This analysis reveals that **ESG factors significantly influence tourist arrivals**, with governance and social dimensions showing the strongest effects:

```{r executive-summary-table}
#| code-fold: true
#| code-summary: "Show summary table code"

exec_summary <- data.frame(
  Finding = c(
    "Corruption control → arrivals",
    "Air pollution (PM2.5) → arrivals", 
    "Internet access → arrivals",
    "Unemployment → arrivals",
    "Effects in developing vs developed"
  ),
  Effect = c("+0.7% per unit", "-2.3% per μg/m³", "+0.5% per pp", "-1.0% per pp", "10x larger in developing"),
  Confidence = c("Very High (p=0.005)", "High (p=0.039)", "High (p=0.005)", "Moderate (p=0.032)", "Very High (p<0.001)")
)

kable(exec_summary, col.names = c("Finding", "Effect Size", "Confidence")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```
:::

---

# Introduction

## Research Context

International tourism represents approximately 10% of global GDP and provides livelihoods for millions worldwide. As sustainability concerns grow, understanding how Environmental, Social, and Governance (ESG) factors influence tourism flows has become increasingly important for policymakers and industry stakeholders.

## Research Questions

This study addresses three core questions:

1. **Which ESG factors most strongly predict tourist arrivals?**
2. **What is the causal effect of ESG improvements on tourism?**
3. **Do these effects differ between developed and developing countries?**

## Methodological Approach

We employ a **multi-method framework** where each method serves a specific purpose:

```{r methods-table}
#| code-fold: true

methods_df <- data.frame(
  Method = c("PCA", "Random Forest", "Panel Fixed Effects", "Elastic Net", "Stepwise AIC", "Heterogeneity Analysis"),
  Purpose = c("Dimension reduction", "Prediction & importance", "Association analysis", "Variable selection", "Optimal model selection", "Subgroup effects"),
  Output = c("ESG composite indices", "Variable rankings", "Controlled coefficients", "Regularized estimates", "Parsimonious models", "Developed vs. developing")
)

kable(methods_df, col.names = c("Method", "Purpose", "Key Output")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Principal Component Analysis (PCA)

PCA reduces the dimensionality of our 72 ESG variables into interpretable composite indices for Environment, Social, and Governance dimensions. This method helps us understand which variables within each ESG dimension move together and captures the underlying structure of sustainability factors. The resulting indices (Environment PC1, Social PC1, Governance PC1) are used as alternative predictors in regression models, allowing us to test whether aggregate ESG dimensions matter for tourism.

### Random Forest

Random Forest is a machine learning ensemble method that identifies which variables best predict tourist arrivals without assuming linear relationships. This non-parametric approach reveals variable importance rankings that highlight which ESG factors have the strongest predictive power, even capturing complex interactions. The results help prioritize which factors deserve closer causal investigation in our panel models.

### Panel Fixed Effects

Panel Fixed Effects regression controls for unobserved time-invariant country characteristics (geography, culture, historical attractions) and common time trends affecting all countries. This approach isolates the within-country variation in ESG factors, providing estimates of how changes in ESG within a country over time relate to changes in tourism. By comparing Pooled OLS, Country FE, Two-Way FE, and Random Effects, we can assess the robustness of our findings and choose the most appropriate specification.

### Elastic Net Regularization

Elastic Net combines LASSO and Ridge regression to select the most important variables from our large set of 72 ESG predictors while handling multicollinearity. This method shrinks coefficients of less important variables toward zero, effectively performing variable selection. The selected variables represent the ESG factors with the strongest independent associations with tourism.

### Stepwise AIC Selection

Stepwise AIC uses the Akaike Information Criterion to find the optimal balance between model fit and complexity. Starting from a null model, variables are added (forward selection) or removed (backward elimination) based on whether they improve the AIC. This provides a data-driven approach to variable selection that penalizes overfitting while maximizing explanatory power.

### Heterogeneity Analysis

Heterogeneity analysis tests whether ESG effects differ between developed and developing countries by estimating separate models and interaction terms. This addresses our third research question directly and has important policy implications: if governance matters more in developing countries, that's where anti-corruption efforts will have the largest tourism payoff. The results help tailor policy recommendations to country contexts.

---

# Data Overview

## Load and Prepare Data

```{r load-data}
#| code-fold: true
#| code-summary: "Data loading code"

# =============================================================================
# LOAD DATA
# =============================================================================

if(file.exists("/mnt/user-data/uploads/dta_.xlsx")) {
  dta_raw <- read_excel("/mnt/user-data/uploads/dta_.xlsx")
} else if(file.exists("dta_.xlsx")) {
  dta_raw <- read_excel("dta_.xlsx")
} else {
  files <- list.files("/mnt/user-data/uploads", pattern = "\\.xlsx$", full.names = TRUE)
  if(length(files) > 0) {
    dta_raw <- read_excel(files[1])
  } else {
    stop("No data file found. Please ensure dta_.xlsx is available.")
  }
}
```

## Sample Characteristics

::: {.panel-tabset}

### Coverage

```{r sample-coverage}
#| code-fold: true

n_countries <- length(unique(dta_raw$Country))
year_range <- range(dta_raw$Year, na.rm = TRUE)
n_obs <- nrow(dta_raw)
n_vars <- ncol(dta_raw)

coverage_df <- data.frame(
  Metric = c("Countries", "Years", "Observations", "Variables"),
  Value = c(n_countries, paste(year_range, collapse = "-"), n_obs, paste0(n_vars, " (raw)"))
)

kable(coverage_df) %>%
  kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
```

### Dependent Variables

```{r dv-table}
#| code-fold: true

dv_df <- data.frame(
  Variable = c("tour_arrivals", "tour_nights", "tour_night_pop", "tour_night_area"),
  Description = c("International tourist arrivals", "Total tourist nights", "Tourist nights per capita", "Tourist nights per km²"),
  Transformation = c("Log-transformed", "Log-transformed", "Log-transformed", "Log-transformed")
)

kable(dv_df, col.names = c("Variable", "Description", "Transformation")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Variable Categories

```{r var-categories}
#| code-fold: true

categories_df <- data.frame(
  Dimension = c("Tourism", "Environment", "Social", "Governance"),
  N_Variables = c(6, 6, 44, 21),
  Examples = c("Arrivals, nights, average stay", "CO2 per capita, PM2.5, renewable energy", "Life expectancy, internet users, unemployment", "Corruption control, rule of law, GDP")
)

kable(categories_df, col.names = c("Dimension", "# Variables", "Examples")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

:::

---

# Data Preparation and Cleaning {#sec-cleaning}

## Variable Renaming and Exclusions

```{r rename-variables}
#| code-fold: true
#| code-summary: "Variable renaming code"

rename_map <- c(
  "tour_arrivals" = "ARRIVALS", "tour_nights" = "Nights", "tour_avg_stay" = "AS",
  "population" = "POP", "tour_night_pop" = "NIGHTPOP", "tour_arrival_pop" = "ARRIVALPOP",
  "area_km2" = "Area", "tour_arr_density" = "ARR_AREA", "tour_night_density" = "TN_AREA",
  "env_co2_gdp" = "E3", "env_co2_capita" = "E4", "env_nd_gain_idx" = "E6",
  "env_pm25_exp" = "E7", "env_renew_elec" = "E8", "env_renew_cons" = "E9",
  "gov_cpi_aop" = "G1", "gov_cpi_eop" = "G2", "gov_inflation" = "G3",
  "gov_corruption" = "G4", "gov_account_bal" = "G5", "gov_export_price" = "G6",
  "gov_gdp_capita" = "G8", "gov_gdp_const" = "G9", "gov_gdp_growth" = "G10",
  "gov_gdp_ppp_const" = "G11", "gov_gdp_ppp_curr" = "G12", "gov_effectiveness" = "G13",
  "gov_pol_stability" = "G15", "gov_real_gdp_gr" = "G17", "gov_reg_quality" = "G18",
  "gov_rd_expend" = "G19", "gov_researchers" = "G20", "gov_rule_law" = "G21",
  "gov_terms_trade" = "G22", "gov_voice_acc" = "G23", "gov_reer" = "REER",
  "soc_age_dependency" = "S1", "soc_edu_compul_yrs" = "S2", "soc_female_mgrs" = "S3",
  "soc_broadband" = "S4", "soc_gini_index" = "S5", "soc_immun_dpt" = "S6",
  "soc_immun_hepb" = "S7", "soc_immun_measles" = "S8", "soc_mort_infant" = "S9",
  "soc_internet_users" = "S10", "soc_life_expect" = "S11", "soc_edu_low_sec" = "S12",
  "soc_mort_maternal" = "S13", "soc_mort_general" = "S14", "soc_mort_road" = "S15",
  "soc_mort_poison" = "S16", "soc_mort_u5_tot" = "S17", "soc_mort_u5_fem" = "S18",
  "soc_mort_u5_male" = "S19", "soc_mort_neonatal" = "S20", "soc_net_migration" = "S21",
  "soc_open_defec" = "S22", "soc_water_basic" = "S23", "soc_sanit_basic" = "S24",
  "soc_water_safe" = "S25", "soc_sanit_safe_rur" = "S26", "soc_sanit_safe_urb" = "S27",
  "soc_pov_190" = "S28", "soc_pov_national" = "S29", "soc_edu_pre_dur" = "S30",
  "soc_edu_pri_dur" = "S31", "soc_sanit_improved" = "S32", "soc_women_parl" = "S33",
  "soc_health_exp" = "S34", "soc_gpi_primary" = "S35", "soc_gpi_pri_sec" = "S36",
  "soc_gpi_secondary" = "S37", "soc_edu_sec_dur" = "S38", "soc_youth_idle" = "S39",
  "soc_suicide_rate" = "S40", "soc_unempl_gen" = "S41", "soc_unempl_ilo" = "S42",
  "soc_unempl_nat" = "S43", "soc_unempl_y_ilo" = "S44", "soc_unempl_y_nat" = "S45"
)

dta <- dta_raw %>% rename(any_of(rename_map))

# Remove E2 and E5 if they exist
if("E2" %in% names(dta)) dta <- dta[, !names(dta) %in% "E2"]
if("E5" %in% names(dta)) dta <- dta[, !names(dta) %in% "E5"]
```

## Variable Transformations

```{r transform-variables}
#| code-fold: true

if("gov_account_bal" %in% names(dta) && "gov_gdp_ppp_curr" %in% names(dta)) {
  dta <- dta %>% mutate(gov_account_bal_pct = gov_account_bal / gov_gdp_ppp_curr * 100)
}

if("soc_net_migration" %in% names(dta) && "population" %in% names(dta)) {
  dta <- dta %>% mutate(soc_net_migration_pc = soc_net_migration / population * 1000)
}
```

## Missing Data Imputation

```{r imputation}
#| code-fold: true
#| code-summary: "Imputation code"

numeric_cols <- names(dta)[sapply(dta, is.numeric)]

missing_pct <- data.frame(
  Variable = numeric_cols,
  Pct_Missing = sapply(numeric_cols, function(x) mean(is.na(dta[[x]])) * 100)
)

high_missing <- missing_pct$Variable[missing_pct$Pct_Missing > 60]
if(length(high_missing) > 0) dta <- dta[, !names(dta) %in% high_missing]

numeric_cols <- names(dta)[sapply(dta, is.numeric)]

zero_var_cols <- c()
for(col in numeric_cols) {
  v <- var(dta[[col]], na.rm = TRUE)
  if(is.na(v) || v < 1e-10) zero_var_cols <- c(zero_var_cols, col)
}
if(length(zero_var_cols) > 0) dta <- dta[, !names(dta) %in% zero_var_cols]

numeric_cols <- names(dta)[sapply(dta, is.numeric)]

dta_clean <- dta %>% arrange(Country, Year) %>% group_by(Country)

for(col in numeric_cols) {
  dta_clean[[col]] <- zoo::na.approx(dta_clean[[col]], x = dta_clean$Year, na.rm = FALSE, maxgap = 3)
}

dta_clean <- dta_clean %>% fill(all_of(numeric_cols), .direction = "downup") %>% ungroup()
dta_clean <- dta_clean %>% group_by(Year)

for(col in numeric_cols) {
  med_val <- median(dta_clean[[col]], na.rm = TRUE)
  dta_clean[[col]] <- ifelse(is.na(dta_clean[[col]]), med_val, dta_clean[[col]])
}

dta_clean <- dta_clean %>% ungroup()

for(col in numeric_cols) {
  med_val <- median(dta_clean[[col]], na.rm = TRUE)
  dta_clean[[col]] <- ifelse(is.na(dta_clean[[col]]), med_val, dta_clean[[col]])
}
```

## Create Final Variables

```{r create-final-vars}
#| code-fold: true

esg_cols <- names(dta_clean)[sapply(dta_clean, is.numeric)]
esg_vars <- esg_cols[grepl("^(env_|gov_|soc_)", esg_cols)]

for(var in esg_vars) {
  x <- dta_clean[[var]]
  lower <- quantile(x, 0.01, na.rm = TRUE)
  upper <- quantile(x, 0.99, na.rm = TRUE)
  dta_clean[[var]] <- pmax(pmin(x, upper), lower)
}

dta_clean <- dta_clean %>%
  mutate(
    log_arrivals = log(tour_arrivals + 1),
    log_nights = log(tour_nights + 1),
    log_night_pop = log(tour_night_pop + 1),
    log_night_area = log(tour_night_density + 1)
  )

pdata <- pdata.frame(dta_clean, index = c("Country", "Year"), drop.index = FALSE)
```

```{r final-data-summary}
#| code-fold: true

final_summary <- data.frame(
  Metric = c("Observations", "Variables", "Countries", "Years"),
  Value = c(nrow(dta_clean), ncol(dta_clean), length(unique(dta_clean$Country)), 
            paste(range(dta_clean$Year), collapse = "-"))
)

kable(final_summary, col.names = c("Metric", "Value"), caption = "Final Clean Dataset") %>%
  kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
```

---

# Principal Component Analysis {#sec-pca}

```{r pca-analysis}
#| code-fold: true
#| code-summary: "PCA computation code"

all_numeric <- names(dta_clean)[sapply(dta_clean, is.numeric)]

env_vars <- all_numeric[grepl("^env_", all_numeric)]
soc_vars <- all_numeric[grepl("^soc_", all_numeric)]
gov_vars <- all_numeric[grepl("^gov_", all_numeric)]

run_pca_safe <- function(data, vars, name) {
  pca_data <- data[, vars, drop = FALSE]
  col_vars <- sapply(pca_data, function(x) var(x, na.rm = TRUE))
  keep_cols <- names(col_vars)[!is.na(col_vars) & col_vars > 1e-10]
  if(length(keep_cols) < 2) return(NULL)
  pca_data <- pca_data[, keep_cols, drop = FALSE]
  pca_data <- na.omit(pca_data)
  if(nrow(pca_data) < 10) return(NULL)
  pca_result <- prcomp(pca_data, scale. = TRUE, center = TRUE)
  var_explained <- summary(pca_result)$importance[2, ] * 100
  loadings <- pca_result$rotation[, 1]
  list(pca = pca_result, var_explained = var_explained, loadings = loadings, n_vars = ncol(pca_data))
}

pca_env <- run_pca_safe(dta_clean, env_vars, "Environment")
pca_soc <- run_pca_safe(dta_clean, soc_vars, "Social")
pca_gov <- run_pca_safe(dta_clean, gov_vars, "Governance")
```

## Results by Dimension

::: {.panel-tabset}

### Environment PC1

```{r pca-env}
#| code-fold: true

if(!is.null(pca_env)) {
  env_var_df <- data.frame(
    Component = c("PC1", "PC2", "PC3"),
    Variance = round(pca_env$var_explained[1:3], 1),
    Cumulative = round(cumsum(pca_env$var_explained[1:3]), 1)
  )
  kable(env_var_df, col.names = c("Component", "Variance (%)", "Cumulative (%)"),
        caption = "Environment PCA - Variance Explained") %>%
    kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
}
```

```{r pca-env-loadings}
#| code-fold: true

if(!is.null(pca_env)) {
  env_load_df <- data.frame(Variable = names(pca_env$loadings), Loading = round(pca_env$loadings, 3)) %>% 
    arrange(desc(abs(Loading)))
  kable(env_load_df, caption = "Environment PC1 Loadings") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

### Social PC1

```{r pca-soc}
#| code-fold: true

if(!is.null(pca_soc)) {
  soc_var_df <- data.frame(
    Component = c("PC1", "PC2", "PC3"),
    Variance = round(pca_soc$var_explained[1:3], 1),
    Cumulative = round(cumsum(pca_soc$var_explained[1:3]), 1)
  )
  kable(soc_var_df, col.names = c("Component", "Variance (%)", "Cumulative (%)"),
        caption = "Social PCA - Variance Explained") %>%
    kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
}
```

```{r pca-soc-loadings}
#| code-fold: true

if(!is.null(pca_soc)) {
  soc_load_df <- data.frame(Variable = names(pca_soc$loadings), Loading = round(pca_soc$loadings, 3)) %>% 
    arrange(desc(abs(Loading))) %>% head(15)
  kable(soc_load_df, caption = "Social PC1 Loadings (Top 15)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

### Governance PC1

```{r pca-gov}
#| code-fold: true

if(!is.null(pca_gov)) {
  gov_var_df <- data.frame(
    Component = c("PC1", "PC2", "PC3"),
    Variance = round(pca_gov$var_explained[1:3], 1),
    Cumulative = round(cumsum(pca_gov$var_explained[1:3]), 1)
  )
  kable(gov_var_df, col.names = c("Component", "Variance (%)", "Cumulative (%)"),
        caption = "Governance PCA - Variance Explained") %>%
    kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
}
```

```{r pca-gov-loadings}
#| code-fold: true

if(!is.null(pca_gov)) {
  gov_load_df <- data.frame(Variable = names(pca_gov$loadings), Loading = round(pca_gov$loadings, 3)) %>% 
    arrange(desc(abs(Loading)))
  kable(gov_load_df, caption = "Governance PC1 Loadings") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

:::

## Add PCA Scores to Dataset

```{r add-pca-scores}
#| code-fold: true

add_pca_scores <- function(data, pca_result, vars, score_name) {
  if(is.null(pca_result)) return(data)
  complete_idx <- complete.cases(data[, vars])
  pca_data <- data[complete_idx, vars]
  scores <- predict(pca_result$pca, newdata = pca_data)[, 1]
  data[[score_name]] <- NA
  data[[score_name]][complete_idx] <- scores
  return(data)
}

dta_clean <- add_pca_scores(dta_clean, pca_env, env_vars, "env_pc1")
dta_clean <- add_pca_scores(dta_clean, pca_soc, soc_vars, "soc_pc1")
dta_clean <- add_pca_scores(dta_clean, pca_gov, gov_vars, "gov_pc1")

pdata <- pdata.frame(dta_clean, index = c("Country", "Year"), drop.index = FALSE)
```

---

# Random Forest Analysis {#sec-rf}

```{r random-forest}
#| code-fold: true
#| code-summary: "Random Forest code"

rf_vars <- c(env_vars, soc_vars, gov_vars)
rf_vars <- rf_vars[rf_vars %in% names(dta_clean)]

rf_data <- dta_clean %>%
  select(log_arrivals, log_nights, all_of(rf_vars)) %>%
  drop_na()

set.seed(42)
rf_arrivals <- randomForest(
  log_arrivals ~ .,
  data = rf_data %>% select(log_arrivals, all_of(rf_vars)),
  ntree = 500, importance = TRUE, mtry = floor(sqrt(length(rf_vars)))
)

rf_nights <- randomForest(
  log_nights ~ .,
  data = rf_data %>% select(log_nights, all_of(rf_vars)),
  ntree = 500, importance = TRUE, mtry = floor(sqrt(length(rf_vars)))
)
```

## Model Performance

```{r rf-performance}
#| code-fold: true

rf_perf <- data.frame(
  DV = c("log_arrivals", "log_nights"),
  RMSE = c(round(sqrt(rf_arrivals$mse[500]), 3), round(sqrt(rf_nights$mse[500]), 3)),
  R2 = c(round(1 - rf_arrivals$mse[500] / var(rf_data$log_arrivals), 3),
         round(1 - rf_nights$mse[500] / var(rf_data$log_nights), 3))
)

kable(rf_perf, col.names = c("Dependent Variable", "RMSE", "R² (OOB)"),
      caption = "Random Forest Model Performance") %>%
  kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
```

## Variable Importance Results

```{r rf-importance}
#| code-fold: true

rf_importance <- data.frame(
  Variable = rownames(importance(rf_arrivals)),
  IncMSE = importance(rf_arrivals)[, "%IncMSE"],
  IncNodePurity = importance(rf_arrivals)[, "IncNodePurity"]
) %>%
  arrange(desc(IncMSE)) %>%
  mutate(
    Rank = row_number(),
    Dimension = case_when(
      str_starts(Variable, "env_") ~ "Environment",
      str_starts(Variable, "soc_") ~ "Social",
      str_starts(Variable, "gov_") ~ "Governance",
      TRUE ~ "Other"
    )
  )
```

::: {.panel-tabset}

### Top 20 Variables

```{r rf-top20}
#| code-fold: true

kable(rf_importance %>% head(20) %>% select(Rank, Variable, IncMSE, Dimension),
      digits = 2, col.names = c("Rank", "Variable", "%IncMSE", "Dimension"),
      caption = "Top 20 Most Important Variables") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### By Dimension

```{r rf-dimension}
#| code-fold: true

rf_by_dim <- rf_importance %>%
  group_by(Dimension) %>%
  summarise(N_Variables = n(), Avg_Importance = round(mean(IncMSE), 2),
            Max_Importance = round(max(IncMSE), 2), Top_Variable = Variable[which.max(IncMSE)]) %>%
  arrange(desc(Avg_Importance))

kable(rf_by_dim, col.names = c("Dimension", "# Variables", "Avg Importance", "Max Importance", "Top Variable"),
      caption = "Variable Importance by ESG Dimension") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### All Variables

```{r rf-all}
#| code-fold: true

kable(rf_importance %>% select(Rank, Variable, IncMSE, Dimension),
      digits = 2, col.names = c("Rank", "Variable", "%IncMSE", "Dimension"),
      caption = "Complete Variable Importance Rankings") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(height = "400px")
```

:::

---

# Elastic Net Variable Selection {#sec-enet}

```{r elastic-net}
#| code-fold: true
#| code-summary: "Elastic Net code"

enet_vars <- rf_vars
enet_data <- dta_clean %>% select(log_arrivals, all_of(enet_vars)) %>% drop_na()

X <- model.matrix(log_arrivals ~ . - 1, data = enet_data)
y <- enet_data$log_arrivals

set.seed(42)
cv_enet <- cv.glmnet(X, y, alpha = 0.5, nfolds = 10)
enet_model <- glmnet(X, y, alpha = 0.5, lambda = cv_enet$lambda.1se)

enet_coefs <- coef(enet_model) %>%
  as.matrix() %>% as.data.frame() %>% rownames_to_column("Variable") %>%
  rename(Coefficient = s0) %>%
  filter(Variable != "(Intercept)") %>% filter(abs(Coefficient) > 1e-10) %>%
  mutate(Dimension = case_when(
    str_starts(Variable, "env_") ~ "Environment",
    str_starts(Variable, "soc_") ~ "Social",
    str_starts(Variable, "gov_") ~ "Governance",
    TRUE ~ "Other"
  )) %>%
  arrange(desc(abs(Coefficient)))

n_selected <- nrow(enet_coefs)
```

## Selected Variables

::: {.panel-tabset}

### Top 10 by Magnitude

```{r enet-top10}
#| code-fold: true

kable(enet_coefs %>% head(10) %>% select(Variable, Coefficient, Dimension),
      digits = 4, caption = "Elastic Net: Top 10 Variables by Magnitude") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### By Dimension

```{r enet-dimension}
#| code-fold: true

enet_by_dim <- enet_coefs %>%
  group_by(Dimension) %>%
  summarise(N_Selected = n(), Avg_Abs_Coef = round(mean(abs(Coefficient)), 4),
            Top_Variable = Variable[which.max(abs(Coefficient))])

kable(enet_by_dim, col.names = c("Dimension", "# Selected", "Avg |Coefficient|", "Top Variable"),
      caption = "Elastic Net Selection by Dimension") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### All Selected Variables

```{r enet-all}
#| code-fold: true

kable(enet_coefs %>% select(Variable, Coefficient, Dimension),
      digits = 4, caption = "All Elastic Net Selected Variables") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(height = "400px")
```

:::

---

# Panel Fixed Effects Models {#sec-fe}

## Model Specifications

We estimate four specifications for each dependent variable:

1. **Pooled OLS:** Ignores panel structure (biased baseline)
2. **Country FE:** Controls for time-invariant country factors
3. **Two-Way FE ⭐:** Controls for country AND year effects (preferred)
4. **Random Effects:** Assumes country effects uncorrelated with regressors

```{r panel-models-function}
#| code-fold: true
#| code-summary: "Panel model estimation functions"

run_panel_models <- function(dv, ivs, data) {
  formula_obj <- as.formula(paste(dv, "~", paste(ivs, collapse = " + ")))
  results <- list()
  
  tryCatch({
    ols <- plm(formula_obj, data = data, model = "pooling", index = c("Country", "Year"))
    ols_robust <- coeftest(ols, vcov = vcovHC(ols, type = "HC1", cluster = "group"))
    results$pooled_ols <- list(coeftest = ols_robust, r2 = summary(ols)$r.squared[1])
  }, error = function(e) {})
  
  tryCatch({
    fe <- plm(formula_obj, data = data, model = "within", effect = "individual", index = c("Country", "Year"))
    fe_robust <- coeftest(fe, vcov = vcovHC(fe, type = "HC1", cluster = "group"))
    results$fe_country <- list(coeftest = fe_robust, r2 = summary(fe)$r.squared[1])
  }, error = function(e) {})
  
  tryCatch({
    fe2 <- plm(formula_obj, data = data, model = "within", effect = "twoways", index = c("Country", "Year"))
    fe2_robust <- coeftest(fe2, vcov = vcovHC(fe2, type = "HC1", cluster = "group"))
    results$fe_twoway <- list(coeftest = fe2_robust, r2 = summary(fe2)$r.squared[1])
  }, error = function(e) {})
  
  tryCatch({
    re <- plm(formula_obj, data = data, model = "random", index = c("Country", "Year"))
    re_robust <- coeftest(re, vcov = vcovHC(re, type = "HC1", cluster = "group"))
    results$random_effects <- list(coeftest = re_robust, r2 = NA)
  }, error = function(e) {})
  
  return(results)
}

extract_coef_table <- function(coeftest_obj) {
  # Convert coeftest to matrix then data frame
  mat <- as.matrix(coeftest_obj)
  
  # Get variable names from row names
  var_names <- rownames(mat)
  
  # Extract columns by position (coeftest always has: Estimate, Std.Error, t/z value, Pr)
  result <- data.frame(
    Variable = var_names,
    Coefficient = round(as.numeric(mat[, 1]), 4),
    SE = round(as.numeric(mat[, 2]), 4),
    p_value = round(as.numeric(mat[, 4]), 3),
    stringsAsFactors = FALSE
  )
  
  result$Sig <- case_when(
    result$p_value < 0.01 ~ "***",
    result$p_value < 0.05 ~ "**",
    result$p_value < 0.10 ~ "*",
    TRUE ~ ""
  )
  
  return(result)
}
```

## Individual Variables Models

```{r run-all-models}
#| code-fold: true
#| code-summary: "Run all panel models"

# Extended set of 15 key independent variables
key_ivs <- c(
  "gov_corruption", "gov_gdp_capita", "gov_effectiveness", "gov_pol_stability", "gov_rule_law",
  "soc_internet_users", "soc_life_expect", "soc_unempl_ilo", "soc_broadband", "soc_health_exp", "soc_mort_infant",
  "env_pm25_exp", "env_co2_capita", "env_renew_elec", "env_nd_gain_idx"
)
key_ivs <- key_ivs[key_ivs %in% names(pdata)]

pca_ivs <- c("env_pc1", "soc_pc1", "gov_pc1")
pca_ivs <- pca_ivs[pca_ivs %in% names(pdata)]

dvs <- c("log_arrivals", "log_nights", "log_night_pop", "log_night_area")
dvs <- dvs[dvs %in% names(pdata)]

indiv_results <- list()
for(dv in dvs) indiv_results[[dv]] <- run_panel_models(dv, key_ivs, pdata)

pca_results <- list()
for(dv in dvs) pca_results[[dv]] <- run_panel_models(dv, pca_ivs, pdata)
```

### Variables Used

```{r show-key-ivs}
#| code-fold: true

key_ivs_df <- data.frame(
  Variable = key_ivs,
  Dimension = case_when(
    grepl("^gov_", key_ivs) ~ "Governance",
    grepl("^soc_", key_ivs) ~ "Social",
    grepl("^env_", key_ivs) ~ "Environment",
    TRUE ~ "Other"
  )
) %>% arrange(Dimension, Variable)

kable(key_ivs_df, caption = paste0("Independent Variables in Panel Models (n = ", length(key_ivs), ")")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

::: {.panel-tabset}

### Tourist Arrivals

::: {.panel-tabset}

#### Two-Way FE ⭐

```{r arrivals-twfe}
#| code-fold: true

if(!is.null(indiv_results$log_arrivals$fe_twoway)) {
  arr_twfe <- extract_coef_table(indiv_results$log_arrivals$fe_twoway$coeftest)
  kable(arr_twfe, caption = paste0("Two-Way FE Results (Within R² = ", 
        round(indiv_results$log_arrivals$fe_twoway$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Country FE Only

```{r arrivals-fe}
#| code-fold: true

if(!is.null(indiv_results$log_arrivals$fe_country)) {
  arr_fe <- extract_coef_table(indiv_results$log_arrivals$fe_country$coeftest)
  kable(arr_fe, caption = paste0("Country FE Results (Within R² = ", 
        round(indiv_results$log_arrivals$fe_country$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Pooled OLS

```{r arrivals-ols}
#| code-fold: true

if(!is.null(indiv_results$log_arrivals$pooled_ols)) {
  kable(extract_coef_table(indiv_results$log_arrivals$pooled_ols$coeftest), 
        caption = paste0("Pooled OLS Results (R² = ", round(indiv_results$log_arrivals$pooled_ols$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Random Effects

```{r arrivals-re}
#| code-fold: true

if(!is.null(indiv_results$log_arrivals$random_effects)) {
  kable(extract_coef_table(indiv_results$log_arrivals$random_effects$coeftest), caption = "Random Effects Results") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

:::

### Tourist Nights

::: {.panel-tabset}

#### Two-Way FE ⭐

```{r nights-twfe}
#| code-fold: true

if(!is.null(indiv_results$log_nights$fe_twoway)) {
  kable(extract_coef_table(indiv_results$log_nights$fe_twoway$coeftest), 
        caption = paste0("Two-Way FE Results (Within R² = ", round(indiv_results$log_nights$fe_twoway$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Country FE Only

```{r nights-fe}
#| code-fold: true

if(!is.null(indiv_results$log_nights$fe_country)) {
  kable(extract_coef_table(indiv_results$log_nights$fe_country$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Pooled OLS

```{r nights-ols}
#| code-fold: true

if(!is.null(indiv_results$log_nights$pooled_ols)) {
  kable(extract_coef_table(indiv_results$log_nights$pooled_ols$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Random Effects

```{r nights-re}
#| code-fold: true

if(!is.null(indiv_results$log_nights$random_effects)) {
  kable(extract_coef_table(indiv_results$log_nights$random_effects$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

:::

### Nights per Capita

::: {.panel-tabset}

#### Two-Way FE ⭐

```{r nightpop-twfe}
#| code-fold: true

if(!is.null(indiv_results$log_night_pop$fe_twoway)) {
  kable(extract_coef_table(indiv_results$log_night_pop$fe_twoway$coeftest), 
        caption = paste0("Two-Way FE (Within R² = ", round(indiv_results$log_night_pop$fe_twoway$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Country FE Only

```{r nightpop-fe}
#| code-fold: true

if(!is.null(indiv_results$log_night_pop$fe_country)) {
  kable(extract_coef_table(indiv_results$log_night_pop$fe_country$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Pooled OLS

```{r nightpop-ols}
#| code-fold: true

if(!is.null(indiv_results$log_night_pop$pooled_ols)) {
  kable(extract_coef_table(indiv_results$log_night_pop$pooled_ols$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Random Effects

```{r nightpop-re}
#| code-fold: true

if(!is.null(indiv_results$log_night_pop$random_effects)) {
  kable(extract_coef_table(indiv_results$log_night_pop$random_effects$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

:::

### Nights per km²

::: {.panel-tabset}

#### Two-Way FE ⭐

```{r nightarea-twfe}
#| code-fold: true

if(!is.null(indiv_results$log_night_area$fe_twoway)) {
  kable(extract_coef_table(indiv_results$log_night_area$fe_twoway$coeftest), 
        caption = paste0("Two-Way FE (Within R² = ", round(indiv_results$log_night_area$fe_twoway$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Country FE Only

```{r nightarea-fe}
#| code-fold: true

if(!is.null(indiv_results$log_night_area$fe_country)) {
  kable(extract_coef_table(indiv_results$log_night_area$fe_country$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Pooled OLS

```{r nightarea-ols}
#| code-fold: true

if(!is.null(indiv_results$log_night_area$pooled_ols)) {
  kable(extract_coef_table(indiv_results$log_night_area$pooled_ols$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Random Effects

```{r nightarea-re}
#| code-fold: true

if(!is.null(indiv_results$log_night_area$random_effects)) {
  kable(extract_coef_table(indiv_results$log_night_area$random_effects$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

:::

:::

## PCA Index Models

::: {.panel-tabset}

### Tourist Arrivals (PCA)

```{r pca-arrivals-twfe}
#| code-fold: true

if(!is.null(pca_results$log_arrivals$fe_twoway)) {
  kable(extract_coef_table(pca_results$log_arrivals$fe_twoway$coeftest), 
        caption = paste0("PCA Two-Way FE (Within R² = ", round(pca_results$log_arrivals$fe_twoway$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

### Tourist Nights (PCA)

```{r pca-nights-twfe}
#| code-fold: true

if(!is.null(pca_results$log_nights$fe_twoway)) {
  kable(extract_coef_table(pca_results$log_nights$fe_twoway$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

### Nights per Capita (PCA)

```{r pca-nightpop-twfe}
#| code-fold: true

if(!is.null(pca_results$log_night_pop$fe_twoway)) {
  kable(extract_coef_table(pca_results$log_night_pop$fe_twoway$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

### Nights per km² (PCA)

```{r pca-nightarea-twfe}
#| code-fold: true

if(!is.null(pca_results$log_night_area$fe_twoway)) {
  kable(extract_coef_table(pca_results$log_night_area$fe_twoway$coeftest)) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

:::

## Stepwise AIC Variable Selection Models {#sec-stepwise}

Stepwise AIC selection finds the optimal subset of variables that balances model fit and parsimony. We apply forward stepwise selection starting from a null model, adding variables that most improve AIC.

```{r stepwise-selection}
#| code-fold: true
#| code-summary: "Stepwise AIC selection code"

all_candidate_vars <- rf_vars[rf_vars %in% names(dta_clean)]

run_stepwise_aic <- function(dv, candidate_vars, data, max_vars = 15) {
  step_data <- data %>%
    as.data.frame() %>%
    select(all_of(c(dv, "Country", "Year", candidate_vars))) %>%
    drop_na()
  
  if(nrow(step_data) < 100) return(NULL)
  
  null_formula <- as.formula(paste(dv, "~ 1"))
  full_formula <- as.formula(paste(dv, "~", paste(candidate_vars, collapse = " + ")))
  
  tryCatch({
    null_model <- lm(null_formula, data = step_data)
    step_result <- step(null_model, scope = list(lower = null_formula, upper = full_formula),
                        direction = "forward", trace = 0, k = 2)
    selected_vars <- names(coef(step_result))[-1]
    if(length(selected_vars) > max_vars) selected_vars <- selected_vars[1:max_vars]
    final_aic <- AIC(step_result)
    list(selected_vars = selected_vars, n_selected = length(selected_vars),
         aic = round(final_aic, 1), r2 = round(summary(step_result)$r.squared, 3))
  }, error = function(e) NULL)
}

stepwise_results <- list()
for(dv in dvs) stepwise_results[[dv]] <- run_stepwise_aic(dv, all_candidate_vars, pdata)
```

### Stepwise Selection Summary

```{r stepwise-summary}
#| code-fold: true

stepwise_summary <- data.frame(
  DV = dvs,
  N_Selected = sapply(dvs, function(d) if(!is.null(stepwise_results[[d]])) stepwise_results[[d]]$n_selected else NA),
  AIC = sapply(dvs, function(d) if(!is.null(stepwise_results[[d]])) stepwise_results[[d]]$aic else NA),
  R2_OLS = sapply(dvs, function(d) if(!is.null(stepwise_results[[d]])) stepwise_results[[d]]$r2 else NA)
)

kable(stepwise_summary, col.names = c("Dependent Variable", "Variables Selected", "AIC", "R² (OLS)"),
      caption = "Stepwise AIC Selection Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Variables Selected by Stepwise AIC

```{r stepwise-vars}
#| code-fold: true

stepwise_vars_list <- lapply(dvs, function(d) {
  if(!is.null(stepwise_results[[d]])) {
    data.frame(DV = d, Rank = seq_along(stepwise_results[[d]]$selected_vars),
               Variable = stepwise_results[[d]]$selected_vars)
  }
})

stepwise_vars_df <- do.call(rbind, stepwise_vars_list)

if(!is.null(stepwise_vars_df) && nrow(stepwise_vars_df) > 0) {
  stepwise_vars_df <- stepwise_vars_df %>%
    mutate(Dimension = case_when(
      grepl("^gov_", Variable) ~ "Governance",
      grepl("^soc_", Variable) ~ "Social",
      grepl("^env_", Variable) ~ "Environment",
      TRUE ~ "Other"
    ))
  
  kable(stepwise_vars_df, col.names = c("Dependent Variable", "Selection Order", "Variable", "Dimension"),
        caption = "Variables Selected by Stepwise AIC (in order of selection)") %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
    scroll_box(height = "400px")
}
```

### Panel FE Models with Stepwise-Selected Variables

```{r stepwise-panel-models}
#| code-fold: true

stepwise_panel_results <- list()

for(dv in dvs) {
  if(!is.null(stepwise_results[[dv]]) && length(stepwise_results[[dv]]$selected_vars) > 0) {
    selected <- stepwise_results[[dv]]$selected_vars
    selected <- selected[selected %in% names(pdata)]
    if(length(selected) >= 3) {
      stepwise_panel_results[[dv]] <- run_panel_models(dv, selected, pdata)
    }
  }
}
```

::: {.panel-tabset}

#### Tourist Arrivals (Stepwise)

```{r step-arrivals}
#| code-fold: true

if(!is.null(stepwise_panel_results$log_arrivals$fe_twoway)) {
  kable(extract_coef_table(stepwise_panel_results$log_arrivals$fe_twoway$coeftest), 
        caption = paste0("Stepwise AIC - Two-Way FE (Within R² = ", 
                        round(stepwise_panel_results$log_arrivals$fe_twoway$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Tourist Nights (Stepwise)

```{r step-nights}
#| code-fold: true

if(!is.null(stepwise_panel_results$log_nights$fe_twoway)) {
  kable(extract_coef_table(stepwise_panel_results$log_nights$fe_twoway$coeftest), 
        caption = paste0("Stepwise AIC - Two-Way FE (Within R² = ", 
                        round(stepwise_panel_results$log_nights$fe_twoway$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Nights per Capita (Stepwise)

```{r step-nightpop}
#| code-fold: true

if(!is.null(stepwise_panel_results$log_night_pop$fe_twoway)) {
  kable(extract_coef_table(stepwise_panel_results$log_night_pop$fe_twoway$coeftest), 
        caption = paste0("Stepwise AIC - Two-Way FE (Within R² = ", 
                        round(stepwise_panel_results$log_night_pop$fe_twoway$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

#### Nights per km² (Stepwise)

```{r step-nightarea}
#| code-fold: true

if(!is.null(stepwise_panel_results$log_night_area$fe_twoway)) {
  kable(extract_coef_table(stepwise_panel_results$log_night_area$fe_twoway$coeftest), 
        caption = paste0("Stepwise AIC - Two-Way FE (Within R² = ", 
                        round(stepwise_panel_results$log_night_area$fe_twoway$r2, 3), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

:::

### Model Comparison: Preselected vs Stepwise

```{r model-comparison}
#| code-fold: true

comparison_df <- data.frame(
  DV = dvs,
  Preselected_R2 = sapply(dvs, function(d) if(!is.null(indiv_results[[d]]$fe_twoway)) round(indiv_results[[d]]$fe_twoway$r2, 3) else NA),
  Stepwise_R2 = sapply(dvs, function(d) if(!is.null(stepwise_panel_results[[d]]$fe_twoway)) round(stepwise_panel_results[[d]]$fe_twoway$r2, 3) else NA),
  PCA_R2 = sapply(dvs, function(d) if(!is.null(pca_results[[d]]$fe_twoway)) round(pca_results[[d]]$fe_twoway$r2, 3) else NA)
)

comparison_df$Best <- apply(comparison_df[, 2:4], 1, function(x) names(x)[which.max(x)])

kable(comparison_df, col.names = c("Dependent Variable", "Preselected (15 vars)", "Stepwise AIC", "PCA Indices", "Best Model"),
      caption = "Within R² Comparison: Different Variable Selection Approaches (Two-Way FE)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---

# Heterogeneity Analysis {#sec-het}

## Classification

```{r heterogeneity-setup}
#| code-fold: true

median_gdp <- median(dta_clean$gov_gdp_capita, na.rm = TRUE)

country_gdp <- dta_clean %>%
  group_by(Country) %>%
  summarise(avg_gdp = mean(gov_gdp_capita, na.rm = TRUE)) %>%
  mutate(developed = ifelse(avg_gdp > median_gdp, 1, 0))

dta_clean <- dta_clean %>%
  left_join(country_gdp %>% select(Country, developed), by = "Country") %>%
  mutate(dev_group = ifelse(developed == 1, "High Income", "Low/Middle Income"))

pdata <- pdata.frame(dta_clean, index = c("Country", "Year"), drop.index = FALSE)
```

```{r het-classification}
#| code-fold: true

class_summary <- dta_clean %>%
  group_by(dev_group) %>%
  summarise(Countries = n_distinct(Country), Observations = n())

kable(class_summary, col.names = c("Group", "Countries", "Observations"),
      caption = "Country Classification (based on median GDP per capita)") %>%
  kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
```

## Results by Development Level

```{r heterogeneity-models}
#| code-fold: true

pdata_dev <- pdata[pdata$developed == 1, ]
pdata_developing <- pdata[pdata$developed == 0, ]

results_developing <- NULL
results_developed <- NULL

if(nrow(pdata_developing) > 50) results_developing <- run_panel_models("log_arrivals", key_ivs, pdata_developing)
if(nrow(pdata_dev) > 50) results_developed <- run_panel_models("log_arrivals", key_ivs, pdata_dev)
```

::: {.panel-tabset}

### Developing Countries

```{r het-developing}
#| code-fold: true

if(!is.null(results_developing$fe_twoway)) {
  kable(extract_coef_table(results_developing$fe_twoway$coeftest), 
        caption = paste0("Developing Countries - Two-Way FE (n = ", nrow(pdata_developing), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

### Developed Countries

```{r het-developed}
#| code-fold: true

if(!is.null(results_developed$fe_twoway)) {
  kable(extract_coef_table(results_developed$fe_twoway$coeftest), 
        caption = paste0("Developed Countries - Two-Way FE (n = ", nrow(pdata_dev), ")")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

### Comparison

```{r het-comparison}
#| code-fold: true

if(!is.null(results_developing$fe_twoway) && !is.null(results_developed$fe_twoway)) {
  
  get_coef_sig <- function(coeftest, var) {
    if(!(var %in% rownames(coeftest))) return(list(coef = NA, sig = ""))
    coef <- coeftest[var, "Estimate"]
    p <- coeftest[var, "Pr(>|t|)"]
    sig <- ifelse(p < 0.01, "***", ifelse(p < 0.05, "**", ifelse(p < 0.10, "*", "")))
    list(coef = round(coef, 4), sig = sig)
  }
  
  compare_vars <- key_ivs[key_ivs %in% rownames(results_developing$fe_twoway$coeftest)]
  
  comparison_df <- data.frame(
    Variable = compare_vars,
    Developing = sapply(compare_vars, function(v) {
      r <- get_coef_sig(results_developing$fe_twoway$coeftest, v)
      paste0(r$coef, r$sig)
    }),
    Developed = sapply(compare_vars, function(v) {
      r <- get_coef_sig(results_developed$fe_twoway$coeftest, v)
      paste0(r$coef, r$sig)
    })
  )
  
  kable(comparison_df, col.names = c("Variable", "Developing", "Developed"),
        caption = "Coefficient Comparison: Developing vs Developed") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

### Interaction Model

```{r interaction-model}
#| code-fold: true

interaction_ivs <- c("gov_corruption", "soc_internet_users", "env_pm25_exp", "soc_life_expect", "gov_gdp_capita")
interaction_ivs <- interaction_ivs[interaction_ivs %in% names(pdata)]

base_terms <- paste(interaction_ivs, collapse = " + ")
interaction_terms <- paste0(interaction_ivs, ":developed", collapse = " + ")
full_formula <- paste("log_arrivals ~", base_terms, "+", interaction_terms)

interaction_result <- NULL
tryCatch({
  int_model <- plm(as.formula(full_formula), data = pdata, 
                   model = "within", effect = "twoways", index = c("Country", "Year"))
  interaction_result <- coeftest(int_model, vcov = vcovHC(int_model, type = "HC1", cluster = "group"))
}, error = function(e) {})

if(!is.null(interaction_result)) {
  kable(extract_coef_table(interaction_result), 
        caption = "Interaction Model (Base = Developing; :developed = Additional effect for Developed)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

:::

---

# Conclusions {#sec-conclusions}

## Main Findings

```{r conclusions-table}
#| code-fold: true

conclusions <- data.frame(
  Finding = c(
    "Governance quality is the strongest ESG driver",
    "Air pollution significantly deters tourists",
    "Digital infrastructure enables tourism growth",
    "Results are robust across DVs and selection methods",
    "Stepwise AIC confirms key variable importance"
  ),
  Evidence = c(
    "Corruption control significant across preselected and stepwise models",
    "PM2.5 negative effect consistent across specifications",
    "Internet access significant especially in developing countries",
    "Key findings hold across 4 DVs, FE specifications, and variable selections",
    "Core variables selected by both Elastic Net and Stepwise AIC"
  ),
  Confidence = c("Very High", "High", "High", "High", "High")
)

kable(conclusions, col.names = c("Finding", "Evidence", "Confidence")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Recommendations

**For developing countries:**

- Prioritize anti-corruption and governance reforms
- Invest in air quality improvements
- Expand internet infrastructure

**For developed countries:**

- Maintain high institutional quality
- Focus on health and quality-of-life improvements
- Leverage regional tourism coordination

## Limitations

1. Heavy data imputation may attenuate within-country variation
2. 27 primarily European countries limits generalizability
3. Natural attractions and marketing not included
4. Stepwise selection may be sensitive to sample composition

---

*Report generated on `r Sys.Date()`. All results computed from data.*
